---
title: "SCC.461 Final Project"
author: "Rawan Abdulelsadig - 35324987"
date: "01/01/2020"
output:
  html_document: default
  pdf_document: default
---

```{r  message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse)
library(gridExtra)
library(car)
```

#### Preparing a paired t-test function for statistical comparisions:
```{r}
Paired.t.test <- function(x1 , x2){
   diffs=x1-x2
   n=length(diffs)
   
   t=mean(diffs)/((sd(diffs)/sqrt(n))) # the t-statistic
   
   print('Mean difference:') 
   print(mean(diffs)) # the MLE
   
   print('t statistic:')
   print(t)
   
   p = pt(-abs(t) , n - 1, lower.tail = T) # calculating the area under the curve at the t percentile (at the left half)
   cat('The P-value obtained:', 2*p, '\n')
   return(2*p)
}
```

#### Preparing a function that plots the performance measures against Min.Leaf.Split, performes the t-tests and prints the results:
```{r}
StatisticalEvaluations <- function(performance, level = 0.05){
  for (i in seq(2,17,2)){
    # making a new dataframe where one column identifies the name of the variable and another specifies the value of the variable
    # to be able to plot both lines in the same graph
    per_melted <- reshape2::melt(performance[c(1,i,i+1)], id.var='Min.Leaf.Split')
    # plot the two raw values 
    p1 <- ggplot(per_melted, aes(x=Min.Leaf.Split, y=value, col=variable)) +geom_point()+ geom_line(size = 1)
    # plot two regression lines
    p2 <- ggplot(per_melted, aes(x=Min.Leaf.Split, y=value, col=variable))+ geom_smooth(method="auto", se=TRUE, fullrange=FALSE,
                                                                                        level=1-level)
    grid.arrange(p1, p2, nrow=2) # Plotting the two plots in a grid with 2 rows and 1 column
    
    
    cat(names(performance)[i],'\n') # printing the name of the metric
    pval <- Paired.t.test(performance[,i],performance[,i+1]) # performing the t-test and getting the P-value
    cat("Are they significantly different at", level*100, "% level?", "\n")
    cat(pval < level, '\n\n')  # evaluating the p-value using the significance level
    }
  
}
```

### First, the case:
#### Criterion: Gini
#### Split :  Best Split

```{r}
per1 <- read.csv('DT-infogain0.0-best-gini-Results.csv')
per1 = per1[1:100,]
```

```{r message=FALSE, warning=FALSE}
StatisticalEvaluations(per1 , level = 0.05)
```

### Second, the case:
#### Criterion: Gini
#### Split :  Random Split

```{r}
per2 <- read.csv('DT-infogain0.0-random-gini-Results.csv')
```

```{r message=FALSE, warning=FALSE}
StatisticalEvaluations(per2 , level = 0.05)
```

### Third, the case:
#### Criterion: Entropy
#### Split :  Best Split

```{r}
per3 <- read.csv('DT-infogain0.0-best-entropy-Results.csv')
per3 = per3[1:100,]
```

```{r message=FALSE, warning=FALSE}
StatisticalEvaluations(per3 , level = 0.05)
```

### Forth, the case:
#### Criterion: Entropy
#### Split :  Random Split

```{r}
per4 <- read.csv('DT-infogain0.0-random-entropy-Results.csv')
```

```{r message=FALSE, warning=FALSE}
StatisticalEvaluations(per4 , level = 0.05)
```
